---
title: "Exercise 1"
author: "Pham Thi Thai - T00727094"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
# The Iris Flower Dataset
head(iris)
```
```{r}
# Numeric summaries
mean <- lapply(iris[, 1:4], mean)
sd <- lapply(iris[, 1:4], sd)
mean(iris$Sepal.Length)

```

```{r}
cbind (mean, sd)
```
```{r}
summary(iris$Species)
```
```{r}
# Simple graph
plot (iris$Sepal.Width, iris$Sepal.Length)
```
```{r}
# Graph using ggplot2
library(ggplot2)
ggplot(data=iris) + 
geom_point(aes(x = Sepal.Width, y = Sepal.Length, 
               color=Species, shape=Species)) +
xlab("Sepal Width") +  
ylab("Sepal Length") +
ggtitle("Sepal Length-Width")
```
```{r}
# Histogram
library (ggplot2)
ggplot (data = iris) +
  geom_histogram (aes(x = Sepal.Width), color = "blue", 
                  fill = "light blue", bins = 10) +
  xlab("Sepal Width") +  ggtitle("Histogram of Sepal Width")
```
```{r}
# Barchart
library (ggplot2)
ggplot (data = iris) +
  geom_bar (aes(x = Species, fill = Species)) +
  xlab("Species") + ggtitle("Bar Chart of Species")
```

```{r}
# Boxplot
library (ggplot2)
ggplot (data = iris) +
  geom_boxplot (aes(x = Species, y = Sepal.Length, fill = Species)) +
  xlab ("Species") + ylab("Sepal Length") +  
  ggtitle("Boxplot of Sepal Length for Species")
```
```{r}
# Scatterplots
library(ggplot2)
ggplot(data=iris, aes(x = Sepal.Width, y = Sepal.Length, 
                      color=Species, shape=Species)) + 
geom_point() +
geom_smooth(method=lm, se=TRUE, fullrange=TRUE) +
xlab("Sepal Width") + ylab("Sepal Length") +
ggtitle("Sepal Length-Width Scatterplot")

```
```{r}
# Scatterplots
library(ggplot2)
ggplot(data=iris, aes(x = Sepal.Width, y = Sepal.Length, 
                      color=Species, shape=Species)) + 
geom_point() +
geom_smooth(method=lm, se=FALSE, fullrange=TRUE) +
xlab("Sepal Width") + ylab("Sepal Length") +
ggtitle("Sepal Length-Width Scatterplot")
```


```{r}

library(ggplot2)
ggplot(data=iris, aes(x = Sepal.Width, y = Sepal.Length, 
                      color=Species, shape=Species)) + 
geom_point() +
xlab("Sepal Width") + 
ylab("Sepal Length") +
ggtitle("Sepal Length-Width Scatterplot") +
facet_grid(cols = vars (Species))+
geom_smooth()
```
```

```{r}
# Date in R
xd <- as.Date("2012-07-27")
weekdays(xd)
```
```{r}
# Install and load package for financial data
pkg_list = c('quantmod', 'zoo', 'tseries', 'PEIP','gridExtra','moments')
# Install packages if needed
for (pkg in pkg_list)
{
  # Try loading the library.
  if ( ! library(pkg, logical.return=TRUE, character.only=TRUE) )
    {
         # If the library cannot be loaded, install it; then load.
        install.packages(pkg)
        library(pkg, character.only=TRUE)
  }
}
```


```{r}
# Stock Prices
start.date = '2019-12-1' # starting date of stock
end.date = '2022-12-1' # ending date of stock
# Download the selected stocks from Yahoo finance
getSymbols(c('AAPL','MSFT','IBM', "GOOG","AMZN","TSLA"), 
              src = "yahoo", from = start.date, to = end.date)
stocks <- merge(AAPL = AAPL[, "AAPL.Adjusted"], MSFT = MSFT[, "MSFT.Adjusted"], 
                IBM = IBM[, "IBM.Adjusted"], GOOG = GOOG[, "GOOG.Adjusted"],
                AMZN = AMZN[, "AMZN.Adjusted"], TSLA = TSLA[, "TSLA.Adjusted"])

names(stocks )<-c("Price.APPLE","Price.MSFT","Price.IBM","Price.GOOG",
                  "Price.AMZN","Price.TSLA")
```
```{r}
# Data
options(width = 70)
head(stocks)
```
```{r}
tail(stocks)
```
```{r}
nrow(stocks)
```
```{r}
# Plot Stock Prices
plot(stocks, legend.loc=1)
```

```{r}
# Log Returns
return.cc = diff(log(stocks)) #calculate log returns
plot(return.cc, legend.loc=1)
```
```{r}

```


