---
title: "Exercise 2"
author: "Pham Thi Thai - T00727094"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# 1. INTRODUCTION
My name is Pham Thi Thai. I was a dedicated mathematics educator at the high school level with a robust academic foundation. I hold a Master's degree in Mathematics, specializing in Analysis. My research interests are centered around leveraging analytical techniques for data exploration, particularly within the domains of finance and education.

# 2. DATA VISUALIZATION
```{r}
# Loading packages
library(tidyverse)
library(palmerpenguins)
library(ggthemes)
```

```{r}
# The penguins data frame
tibble(penguins)
```

```{r}
# Creating a ggplot
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point()

```
```{r}
# Adding aesthetics and layers
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g, color = species)
) +
  geom_point()
```
```{r}
# Drawing the line of best fit
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g, color = species)
) +
  geom_point() +
  geom_smooth(method = "lm")
```
```{r}
# Drawing the unseperated line of best fit
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point(mapping = aes(color = species)) +
  geom_smooth(method = "lm")
```
```{r}
# Different shapes plot
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point(mapping = aes(color = species, shape = species)) +
  geom_smooth(method = "lm")
```
```{r}
# Enhancing a perfect plot 
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point(aes(color = species, shape = species)) +
  geom_smooth(method = "lm") +
  labs(
    title = "Body mass and flipper length",
    subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
    x = "Flipper length (mm)", y = "Body mass (g)",
    color = "Species", shape = "Species"
  ) +
  scale_color_colorblind()
```
```{r}
# Visualizing distributions
# Barchart with non-ordered levels
ggplot(penguins, aes(x = species)) +
  geom_bar()
```
```{r}
# Barchart with ordered levels
ggplot(penguins, aes(x = fct_infreq(species))) +
  geom_bar()
```
```{r}
# Histogram
ggplot(penguins, aes(x = body_mass_g)) +
  geom_histogram(binwidth = 200)
```
```{r}
# Histograms with different binwidths
ggplot(penguins, aes(x = body_mass_g)) +
  geom_histogram(binwidth = 20)
ggplot(penguins, aes(x = body_mass_g)) +
  geom_histogram(binwidth = 2000)
```
```{r}
# Density plot
ggplot(penguins, aes(x = body_mass_g)) +
  geom_density()
#> Warning: Removed 2 rows containing non-finite values (`stat_density()`).
```
```{r}
# Visualizing relationships
# Boxplot
ggplot(penguins, aes(x = species, y = body_mass_g)) +
  geom_boxplot()
```
```{r}
# Density plots
ggplot(penguins, aes(x = body_mass_g, color = species)) +
  geom_density(linewidth = 0.75)
```
```{r}
# Adding transparency to the filled density curves
ggplot(penguins, aes(x = body_mass_g, color = species, fill = species)) +
  geom_density(alpha = 0.5)
```
```{r}
# Stacked bar plots
# The frequencies of each species of penguins on each island
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar()
```

```{r}
# A relative frequency plot 
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(position = "fill")
```
```{r}
# Scatterplot of three or more variables
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point(aes(color = species, shape = island))
```
```{r}
# Facets (subplots that each display one subset of the data)
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point(aes(color = species, shape = species)) +
  facet_wrap(~island)
```


# 3. STOCK PRICE DATA
```{r}
# Loading the data
mydata <- read.csv("AMZN.csv")
```
```{r}
# Print out the first observations
head(mydata)
```
```{r}
# Print out the last observations
tail(mydata)
```
```{r}
# Convert the 'Date' column to a Date type
mydata$Date <- as.Date(mydata$Date)
```
```{r}
# Create a time series plot
ggplot(mydata, aes(x = Date, y = Close)) +
  geom_line(color="blue") +
  labs(title = "Stock Price Time Series",
       x = "Date",
       y = "Stock Price")
```


